30 years of privacy research for the masses \& \newline 15 years of decentralization attempts

### 2014-03-13 paper.tex
Title fix
Remove duplicate text

@@

	\subsection{Bootstrapping new nodes}
		If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be one starting point where this new node can discover neighbors in the network to connect with. In Tor a directory server can tell the new node what his neighbors are and how to find them \cite{dingledine2004tor}.

		Moving this system to a non peer-to-peer base is difficult. Dingledine et al. stated that this indeed is still an open problem. With decentralized systems there is no central directory server to tell a new node where to locate neighbors and there might be a lot of malicious nodes in the network. Some systems such as Tarzan, MorphMix and Pastry \cite{rowstron2001pastry, rennhard2002introducing} do have made some attempts towards a decentralized system, but have yet to find the final answer.

	\subsection{Incentives in decentralized systems}
		As described in section \ref{ss:tor_disadvantages} people are using more bandwidth than that they donate. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words the incentive data has to be accurate and reliable. Besides the integrity of the data, also the traffic it generates on the network should have minimal impact.

		Rahman \cite{rahman2009survey} proposes several options to build in incentives in a peer-to-peer network. The first thing that is described is the so called \emph{Warm-glow Model}. In this model determines the percentage of free-rides based on the probabilistic population distribution. If the percentage is above a certain threshold, the system will show signs of diminishing marginal returns.

@@

		The third proposal is \emph{Reciprocity-Based Schemes}. Using this approach a peer maintains a behavior history of other peers in the network. These schemes can be based on two somewhat reciprocities: direct reciprocity or indirect reciprocity. The formal are more suitable for longer relationships between peers. The latter is more scalable but they rely on third party and must handle trust issues themselves.

	\subsection{Key exchange}
		When using a decentralized network, using a centralized authority for managing the keys, is not possible. This means that for secure communication, peers have to exchange the keys directly with each other, without a trusted party between them. Diffie-Hellman is a very popular algorithm for exchanging keys between two parties. Diffie-Hellman is for example used in the circuit setup in Onion Routing (see chapter I), however, it is possible for an adversary to manipulate with the keys send between two parties. This makes it possible for the adversary to decrypt all message sent between the two parties.

		Tor currently uses an interactive forward-secret key-exchange protocol called the Tor authentication protocol (TAP) [linkje naar ace]. This protocol uses telescoping which means that the initiator negotiates session keys with each successive hop in the circuit. There are several proposals for more efficient key exchange methods. One of them is ACE, an one-way authenticated key exchange protocol. The authors of this methods claim to have an 46\% efficiency improvement on the side of the client and an efficiency improvement of nearly 19\% on the side of the Onion Routers. ACE requires clients to send one extra element but that doesn't introduce any overhead because this element fits in the unused space in a cell.

@@

		There are three types of file keys in Freenet: the first one is called the Keyword-Signed Key (KSK) which is derived from a short description of the file. Another key is the Signed-Subspace key (SSK) which enables personal namespaces. This key contains a public key and a private key where the private key is used to store the data and the public key is used in the queries of the file. The third type of key is the Content-Hash Key (CHK) which is used for updating and splitting of contents.

		The routing algorithm for storing and retrieving data is dynamically and can adjust to the topology of the network. Each peer has only knowledge about his neighbours. Each request has a Hops-To-Live timer which indicates how many peers the request may traversal. Each peer decrements the timer by one and when the timer reaches zero, the request isn't forwarded any more. Results of queries are being cached in intermediate nodes to reduce the time for a query response. To prevent looping of the requests, each request contains a random identifier. The peers the request travels through, keep track of these identifiers and rejects the request if the request has already been answered by the peer.

@@

		MIX networks typically are high latency and traffic moves slow. Messages will have to be stored on a node until enough messages have arrived to start sending them out in a random order. Often times cover traffic is used to generate enough messages for the network to obfuscate the real communication, which in turn generates a lot of bandwidth overhead. MorphMix has been developed as a low latency and high performance network. As such, it will not hold messages for very long, nor will it use cover traffic.

		In contrast to Tor, MorphMix does not feature a centralized node discovery mechanism. Instead, every node is free to chose a set of next nodes that will be the continuation of an anonymous tunnel. A malicious node could select a colluding node to continue the tunnel and therefor control the entire tunnel. To prevent this, a witness node is appointed to mediate the setup of an anonymous tunnel. Although this makes the possibility of such an attack more difficult, it doesn't make it impossible.

		An important part of MorphMix is the ability to detect malicious tunnels. Tunnels that are set up by well behaving nodes will select the next nodes in the tunnel randomly. Malicious colluding nodes, however, will specifically select nodes that are part of the malicious network to continue the tunnel. This will reveal itself in the fact that the probability of the selection of certain nodes is increased. By using this information it becomes possible, to an extent, to detect malicious tunnels.

@@

		Tapestry is a fundamental component of Oceanstore [linkje naar paper over Oceanstore], a decentralized storage system. Tapestry is also used in systems such as Bayeux and SpamWatch, a decentralized spam-filtering system.

@@

	\subsection{Browser based attacks}
		Traffic analysis can be used to attack the anonymity of a user (a Tor client) browsing the web using Tor \cite{abbott2007browser}. By misusing the exit policy of Tor one can reduce the time required to perform the analysis.

		By performing traffic analysis to compare the unique IDs of the web server and the circuits passing trough the entry node, a user can be identified via matches. A similar system can be set up only using the HTML meta refresh tag. To increase the odds of a user choosing the malicious exit node, one can run the exit node on unpopular ports. There are usually only a few exit nodes running on file sharing ports 4661 to 4666.

	\subsection{Sybil attacks}
		The Sybil attack is an attack where a single attacker represents itself as multiple nodes in a peer-to-peer system. Abusing this, the attacker is able to propagate false assumptions about the network to other nodes.

		First described by Douceur, he \cite{douceur2002sybil} mathematically proves that this attack is always possible without a central authority that certifies the participating nodes in one way or the other, except under what he call "extreme and unrealistic assumptions of resource parity and coordination among entities", or in other words: require all participants to do something expensive (in terms of resources) to identify itself. This must be done within a small enough time frame that an attacker can't do this in sequence, but all nodes must do them in parallel.

		A fully distributed network that implements such a solution is the Bitcoin network [BITCOIN PAPER: http://s.kwma.kr/pdf/Bitcoin/bitcoin.pdf], in which computing power, and not the number of nodes is important for the general network consensus.

	\subsection{Replay Attacks}
		Hier wat over replay attacks

@@

	In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore we looked at the main disadvantages and problems Tor currently has. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

	From this table we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of  attacks.

>>>>

		\subsection{Browser based attacks}
		Traffic analysis can be used to attack the anonymity of a user (a Tor client) browsing the web using Tor \cite{abbott2007browser}. By misusing the exit policy of Tor one can reduce the time required to perform the analysis from $O(nk)$ to $O(n+k)$ where $n$ is the number of exit nodes and $k$ is the number of entry guards.\\

		By performing traffic analysis to compare the unique IDs of the web server and the circuits passing trough the entry node, a user can be identified via matches. A similar system can be set up only using the HTML meta refresh tag. To increase the odds of a user choosing the malicious exit node, one can run the exit node on unpopular ports. There are usually only a few exit nodes running on file sharing ports 4661 to 4666. Since Tor prefers older circuits, using a denial of service attack against the older exit nodes forces Tor into creating a circuit with the malicious exit node.

		\subsection{Sybil attacks}
			Hier wat over sybil attacks

		\subsection{Replay Attacks}
			A replay attack [cite naar replay attack paper] happens when a malicious entry node duplicates cells and send them again. Since Tor uses uses the counter mode of Advanced Encryption Standard (AES-CTR) for encryption and decryption, the counter will be wrong when the duplicated package arrives causing the connection to be torn down.

			Using this, an accomplice exit router can therefore in cooperation with the entry node discover the sender and receiver's relationship. On top of that, this attack can also be used as a denial of service attack. This attack is according to the authors quite challenging to solve and does require some further research.

@@

		As we saw in the previous section, we would like to have a decentralized anonymous network. This isn't an easy task: 15 years of decentralization attempts on the Tor network are behind us. Several well-used decentralized peer-to-peer overlay networks are active such as Gnutella [link], Freenet [link] and Tapestry [link]. These networks will be described in more detail in the next chapter. Among these networks are some overlay anonymous networks such as Tarzan and Torks [link].

		\subsection{Bootstrapping new nodes}
			If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be a starting point where this new node can discover neighbors in the network to connect with. In Tor, a directory server can tell the new node what his neighbors are and where to find them \cite{dingledine2004tor}.

			Moving this system to a non peer-to-peer base is difficult: Dingledine et al. stated that this indeed is still an open problem. With decentralized systems, there is no central directory server to tell a new node where to locate neighbors and there might be a lot of malicious nodes in the network. Some systems such as Tarzan, MorphMix and Pastry \cite{rowstron2001pastry, rennhard2002introducing} are decentralized but they do suffer from performance issues. Some of these issues can be found in the last section when comparing these systems.

		\subsection{Incentives in decentralized systems}
			As described in section \ref{ss:tor_disadvantages}, people are usually using more bandwidth than that they donate. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words, the reputation data has to be accurate and reliable. Besides the integrity of this data, the traffic it generates on the network should have minimal impact on the overall performance.

			Rahman \cite{rahman2009survey} proposes several options to build in incentives in a peer-to-peer network. The first proposal that is described is the so called \emph{Warm-glow Model}. This model determines the percentage of free-rides based on the probabilistic population distribution. If the percentage is above a certain threshold, the system will show signs of diminishing marginal returns.

@@

			The third proposal is \emph{Reciprocity-Based Schemes}. Using this approach, a peer maintains a behavior history of other peers in the network. These schemes can be based on two somewhat reciprocities: direct reciprocity or indirect reciprocity. The formal are more suitable for longer relationships between peers. The latter is more scalable but they rely on third party and must handle trust issues themselves.

		\subsection{Key exchange}
			With a decentralized network, using a centralized authority for managing the keys, is not possible. This means that for secure communication, peers have to exchange the keys directly with each other, without a trusted party between them. Diffie-Hellman is a very popular algorithm for exchanging keys between two parties. Diffie-Hellman is for example used in the circuit setup in Onion Routing (see secion II-D), however, it is possible for an adversary to manipulate with the keys send between two parties, making the protocol vulnerable for a man-in-the-middle attack. This weakness makes it possible for an adversary to decrypt all message sent between the two parties.

			Tor currently uses an interactive forward-secret key-exchange protocol called the \emph{Tor Authentication Protocol} (TAP) [linkje naar ace]. This protocol uses telescoping which means that the initiator negotiates session keys with each successive hop in the circuit. There are several proposals for more efficient key exchange methods. One of them is \emph{ACE}, an one-way authenticated key exchange protocol. The authors of this methods claim to have an 46\% efficiency improvement on the side of the client and an efficiency improvement of nearly 19\% on the side of the Onion Routers. ACE requires clients to send one extra element in the key exchange. This however, does not introduce any overhead because the element fits in the unused space in a cell.

@@

		Many other decentralized peer-to-peer networks have been proposed and some have been implemented [linkjes naar papers]. Examples of fully decentralized networks are Gnutella, Freenet, Pastry, Tapestry, and Tarzan. Each of these decentralized networks will be discussed in the next section. Note that we're only considering fully decentralized networks here. Networks that are partially or hybrid centralized, such as KaZaA or Napster, are not discussed here.[link naar survey p2p?]

@@

			There are three types of file keys in Freenet: the first one is called the \emph{Keyword-Signed Key} (KSK) which is derived from a short description of the file. Another key is the \emph{Signed-Subspace key} (SSK) which enables personal namespaces. This key contains a public key and a private key where the private key is used to store the data and the public key is used in the queries of the file. The third type of key is the \emph{Content-Hash Key} (CHK) which is used for updating and splitting of contents.

			The routing algorithm for storing and retrieving data is dynamically and can adjust to the topology of the network. Each peer has only knowledge about his direct neighbours. Each request has a Hops-To-Live timer which indicates how many peers the request may traversal. Each peer decrements the timer by one and when the timer reaches zero, the request isn't forwarded any more. Results of queries are being cached in intermediate nodes to reduce the time for a query response. To prevent looping of the requests, each request contains a random identifier. The peers the request travels through, keep track of these identifiers and rejects the request if the request has already been answered by the peer.

@@

			Tapestry is a fundamental component of OceanStore [linkje naar paper over OceanStore], a decentralized storage system. Tapestry is also used in systems such as Bayeux and SpamWatch, a decentralized spam-filtering system.

@@

@@

		In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore, we looked at the main disadvantages and problems Tor is currently facing. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

		From this table, we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of attacks.
